### 2025-05-05 Chapters/appendix.tex

\section*{Code}
\begin{verbatim}
import nibabel as nib
import cv2
import gc
import os
import numpy as np
from glob import glob
import tensorflow as tf
from tqdm.auto import tqdm
import keras.backend as K
import matplotlib.pyplot as plt
from Codes.Dataloader import *
from Codes.Evalution_matrix import *
from Codes.own_method import *
from sklearn.preprocessing import MinMaxScaler
from sklearn.model_selection import train_test_split
import pandas as pd

class UNET3D(tf.keras.Model):
    def __init__(self, classes, dropout_rate=0.5, l2_reg=0.01):
        super(UNET3D, self).__init__()
        self.classes = classes
        self.dropout_rate = dropout_rate
        self.l2_reg = l2_reg

    def conv3D(self, x, filters, filter_size, activation='relu'):
        out = tf.keras.layers.Conv3D(
            filters,
            (filter_size, filter_size, filter_size),
            padding="same",
            kernel_regularizer=tf.keras.regularizers.l2(self.l2_reg)
        )(x)
        out = tf.keras.layers.BatchNormalization()(out)
        out = tf.keras.layers.Activation('relu')(out)
        out = tf.keras.layers.Dropout(self.dropout_rate)(out)
        return out

    def upsampling3D(self, x, filters, filter_size, stride=2, 
    activation='relu'):
        up_out = tf.keras.layers.Conv3DTranspose(
            filters, (filter_size, filter_size, filter_size),
            strides=(stride, stride, stride), padding='same'
        )(x)
        out = tf.keras.layers.BatchNormalization()(up_out)
        return out

    def concatenate(self, x1, x2):
        concat = tf.keras.layers.concatenate([x1, x2])
        return concat

    def max_pool3D(self, x, filter_size, stride):
        out = tf.keras.layers.MaxPooling3D(
            (filter_size, filter_size, filter_size), strides=stride
        )(x)
        return out

    def downconv(self, x, filters):
        s1 = self.conv3D(x, filters, 3)
        s2 = self.conv3D(s1, filters, 3)
        return s1

    def upconv(self, x, filters, skip_connection):
        e1 = self.upsampling3D(x, filters, 2)
        concat = self.concatenate(e1, skip_connection)
        conv1 = self.conv3D(concat, filters, 3)
        conv2 = self.conv3D(conv1, filters, 3)
        return conv2

    def call(self):
        inputs = tf.keras.layers.Input(shape=(64, 64, 64, 1))
        d1 = self.downconv(inputs, 32)
        m1 = self.max_pool3D(d1, filter_size=2, stride=2)
        d2 = self.downconv(m1, 64)
        m2 = self.max_pool3D(d2, filter_size=2, stride=2)
        d3 = self.downconv(m2, 128)
        m3 = self.max_pool3D(d3, filter_size=2, stride=2)
        d4 = self.downconv(m3, 256)
        m4 = self.max_pool3D(d4, filter_size=2, stride=2)

        bridge = self.conv3D(m4, 1024, 3, 1)
        bridge = self.conv3D(bridge, 1024, 3, 1)

        u1 = self.upconv(bridge, 256, d4)
        u2 = self.upconv(u1, 128, d3)
        u3 = self.upconv(u2, 64, d2)
        u4 = self.upconv(u3, 32, d1)

        logits = tf.keras.layers.Conv3D(self.classes, (1, 1, 1), 
        padding="same")(u4)
        logits = tf.nn.sigmoid(logits)

        model = tf.keras.Model(inputs=[inputs], outputs=[bridge, logits])
        return model

unet1 = UNET3D(classes=1, dropout_rate=0.2, l2_reg=0.01)

imgepath= "/kaggle/input/gbm-64/GBM/image"
maskpath= "/kaggle/input/gbm-64/GBM/mask"

images_file = os.listdir(imgepath)
masks_file = os.listdir(maskpath)

npy_images_path = [os.path.join(imgepath, img_file) for 
img_file in images_file[:]]
npy_masks_path = [os.path.join(maskpath,mas_file) for 
mas_file in masks_file[:]]
npy_images_path.sort()
npy_masks_path.sort()

metrics = [iou,dice_coef,precision,sensitivity,specificity]

LR = 0.00001
batch_size = 5
epochs = 151

data ={}
data["original_data"] = mini_batches_(npy_images_path[:100],
npy_masks_path[:100],batch_size)

data["model"] = unet1.call()
data["optimizer"] = tf.keras.optimizers.Adam(LR)

test_data_gen = mini_batches_(npy_images_path[100:110], 
npy_masks_path[100:110],batch_size)

plot_dice_losses=[]
plot_total_loss=[]
plot_dice_co=[]

past_model = None

for epoch in range(epochs):
    print(epoch+1,'/',epochs)
    model = data['model']
    optimizer = data["optimizer"]
    # optimizer = optimizer
    model.set_weights(data["model"].get_weights())

    dice = []
    pre = []
    batch_loss = []
    se = []
    spe = []
    io = []
    dice_losses=[]
    contrastive_losses=[]

    for batch_idx, (images, masks) in enumerate(tqdm(data['original_data'])):
        pro2, _ = data['model'](images)
        with tf.GradientTape() as tape:
            pro1 , logits = model(images)
            loss_combined = tversky_loss(y_true=masks, y_pred=logits)

            loss = loss_combined
            batch_loss.append(loss)

            dice_loss_value = dice_loss2(masks, logits)
            contrastive_loss_value = contrastive_loss(prev=pro2, pres=pro1)

            dice_losses.append(dice_loss_value)
            contrastive_losses.append(contrastive_loss_value)

            # metrics
            dice.append(dice_coef(masks, logits))
            pre.append(precision(masks, logits))
            se.append(sensitivity(masks, logits))
            spe.append(specificity(masks, logits))
            io.append(iou(masks, logits))

        grads = tape.gradient(loss, model.trainable_variables)
        gradients, _ = tf.clip_by_global_norm(grads, clip_norm=1.0)
        optimizer.apply_gradients(zip(gradients, model.trainable_variables))

    # local_weight_list.append(model.get_weights())
    batch_loss = np.array(batch_loss).mean()
    dice = np.array(dice).mean()
    pre = np.array(pre).mean()
    se = np.array(se).mean()
    spe = np.array(spe).mean()
    io = np.array(io).mean()

    data['past_model'] = model
    data['model'] = model

    mean_dice_loss = np.mean(dice_losses)
    mean_contrastive_loss = np.mean(contrastive_losses)
    print("\nWeighted Combined Loss: {} | Dice Coeff: {}  |
    \n\n Precision: {} Sensitivity: {} | 
    \n\n Specificity: {} , IOU: {} |"
    .format(batch_loss, dice, pre, se, spe, io))

    plot_dice_losses.append(mean_dice_loss)
    plot_total_loss.append(batch_loss)
    plot_dice_co.append(dice)

    if epoch%10 == 0 and epoch != 0:
        dice_test = []
        pre_test = []
        batch_loss_test=[]
        se_test=[]
        spe_test = []
        io_test = []
        model_test = data["model"]
        for batch_idx, (images, masks) in enumerate(tqdm(test_data_gen)):
            predictions, logits = model(images)

            loss = dice_loss2(masks,logits)
            batch_loss_test.append(loss)

            dice_test.append(dice_coef(masks, logits))
            pre_test.append(precision(masks, logits))
            se_test.append(sensitivity(masks, logits))
            spe_test.append(specificity(masks, logits))
            io_test.append(iou(masks, logits))

        batch_loss_test = np.array(batch_loss_test).mean()
        dice_test = np.array(dice_test).mean()
        pre_test =np.array(pre_test).mean()
        se_test =  np.array(se_test).mean()
        spe_test =  np.array(spe_test).mean()
        io_test =  np.array(io_test).mean()
        print("----------->Test<-----------")
        print("Epoch: {} , Loss: {} , Dice Coeff: {}
        \n\n, Precision: {} Sensitivity: {} 
        \n\n Specificity: {} , IOU {}".format(epoch+1,batch_loss_test,dice_test,
        pre_test,se_test,spe_test,io_test))
        print("----------->Trining Conti<-----------")

plt.figure(figsize=(10, 5))
plt.plot(plot_dice_losses, label='Dice Loss', color='blue')
plt.title('Dice Loss Over Epochs')
plt.xlabel('Epochs')
plt.ylabel('Dice Loss')
plt.legend()
plt.show()

# Plot Weighted Combined Loss
plt.figure(figsize=(10, 5))
plt.plot(plot_total_loss, label='Weighted Combined Loss', color='red')
plt.title('Weighted Combined Loss Over Epochs')
plt.xlabel('Epochs')
plt.ylabel('Weighted Combined Loss')
plt.legend()
plt.show()

\end{verbatim}