### 2025-05-05 Title/abstract.tex

\begin{center}
\section*{\textcolor{internationalkleinblue}{\textbf{{Abstract}}}}
\end{center} 
This project presents \textit{SleepGCN-Transformer}, a hybrid model combining Graph Convolutional Networks (GCN) and Transformer encoders for sleep stage classification. Using the SleepEDF dataset, it incorporates four physiological signals: EEG (Fpz-Cz, Pz-Oz), EMG (submental), and EOG (horizontal). Preprocessing includes 30-second epochs, band-pass filtering (0.3--30 Hz), and graph-based EEG channel representation.

The GCN module captures spatial relationships across EEG channels, while the Transformer encoder models temporal dependencies from graph-level embeddings. Focal Loss addresses class imbalances, and a CosineAnnealingLR scheduler optimizes learning rate decay. Training with the AdamW optimizer for 20 epochs achieves 93.12 \% training and 93.04\% validation accuracy.

Model performance, assessed using precision, recall, and F1-score, demonstrates high efficacy. LIME-based feature importance analysis highlights EMG and EEG Pz-Oz channels as key contributors. This hybrid model exhibits state-of-the-art performance, with future directions focused on enhancing explainability and clinical integration.