### 2025-05-05 Title/acknowledgement.tex

%\begin{center}
%\textbf{\large Acknowledgment}
%\end{center} 
\section*{\textcolor{internationalkleinblue}{\textbf{Acknowledgment}}}
 While my name appears as the sole contributor to the completion of this summer internship, it is important to recognize that the guidance and support of many individuals played a significant role in its success. This internship is the result of a collective effort, shaped by the contributions of numerous people, to whom I am deeply grateful.

 I extend my heartfelt appreciation to my internal guide, Dr. Santosh Kumar Satapathy
Author, for his consistent guidance, encouragement, and insightful suggestions throughout this internship. His belief in my abilities provided the confidence and motivation necessary to complete this project. Dr. Santosh Kumar Satapathy
Author’s support, along with his generous provision of time and access to resources, was instrumental in achieving the objectives of this internship.

 I am also profoundly thankful for the unwavering support of my parents. Their continuous inspiration, moral backing, and blessings have been the cornerstone of my journey. Their understanding and encouragement have been invaluable, and I am truly fortunate to have had their steadfast support throughout this endeavor.
\begin{flushright}
\textbf{Tanmay Rathod}\\
\end{flushright}