### 2025-05-05 Title/approval.tex

\begin{center}
    \textbf{\large Approval Sheet}
\end{center}

This thesis entitled \textbf{\enquote{SleepGCN-Transformer: A Hybrid Graph
Convolutional and Transformer Network for Sleep
Stage Classification}} by \textbf{Tanmay Rathod} is recommended for the degree of \textbf{M.Tech} in \textbf{Artificial Intelligence}.
\vspace{0.8cm}

\hspace{9cm}  
\begin{flushright}   
    \makebox[1.8in]{\hrulefill}\\
    Guide : Dr Santosh Satapathy, \\
    Assistant Professor \\
    Dept. of ICT \\

\end{flushright}
\vspace{0.5cm}

\hspace{9cm}  
\begin{flushright}   
    \makebox[1.8in]{\hrulefill}\\
    Dr Nitin Singh Rajput, \\
    Assistant Professor \\
    Dept. of ICT
\end{flushright}
\vspace{0.5cm}

% \hspace{9cm}\\
\begin{flushright}
    \makebox[1.8in]{\hrulefill}\\
    Dr. Paawan Sharma,\\
    Head Of Department \\ 
    Dept. of ICT
\end{flushright}

\vfill
\begin{flushleft}
    Date: \makebox[1.8in]{\hrulefill}\\
    Place: \makebox[1.8in]{\hrulefill}\\    
\end{flushleft}