### 2025-05-05 Title/publications.tex

%\fancyhead[L]{\fontsize{8}{8}\selectfont\itshape\nouppercase{List of Publication}}
%\thispagestyle{empty}

\section*{\textcolor{internationalkleinblue}{\textbf{Publications on Related Topic}}}
\begin{enumerate}
	\item \textbf{C Chodvadiya}, P Sharma, \textquotedblleft Class-Enhanced Mahalanobis Distance Contrastive (CEMDC) Loss for Brain Tumor Segmentation, \textquotedblright \,\textcolor{internationalkleinblue}{\textbf{Under Review with Professor}}

    \item P Divya, \textbf{C Chodvadiya}, B Raman, YW Chen, CK Mohan, \textquotedblleft FSF3A: Federated Spatial Feature Alignment and Adaptive Aggregation for Heterogeneous Brain Tumor Segmentation, \textquotedblright \,\textcolor{internationalkleinblue}{\textbf{Under Review}} in 27TH EUROPEAN CONFERENCE ON ARTIFICIAL INTELLIGENCE.

\end{enumerate}
\newpage

\section*{\textcolor{internationalkleinblue}{\textbf{Other Publications During M.Tech}}}
\begin{enumerate}
   	\item \textbf{C Chodvadiya}, KG Singh, R Vyas*, \textquotedblleft Detection of Palmprint from Contactless Smartphone-Based Video Hand Dataset, \textquotedblright  NCVPRIPG-CHBGR, 2023.

    \item \textbf{C Chodvadiya*}, V Solanki, KG Singh, \textquotedblleft Intelligent Virtual Worlds: A Survey of the Role of AI in the Metaverse, \textquotedblright  In 2024 3rd IEEE International Conference for Innovation in Technology (INOCON).

    \item KG Singh, \textbf{C Chodvadiya}, C Bhatt, P Shah, A Bruno, \textquotedblleft Seeing in the Dark: A Different Approach to Night Vision Face Detection with Thermal IR Images, \textquotedblright  In CEURWORKSHOP PROCEEDINGS. CEUR-WS, 2023.\\ \textbf{DOI : }\textcolor{internationalkleinblue}{\textbf{\url{https://ceur-ws.org/Vol-3563/paper_1.pdf}}}

    \item KG Singh, \textbf{C Chodvadiya}, S Vyas, H Patel, S Mishra, B Tripathi, \textquotedblleft Prediction of Solar Power Generation and Maintenance Activities for1MWPower Plant, \textquotedblright In 2023 IEEE IAS Global Conference on Emerging Technologies (GlobConET), pp. 1-6. IEEE, 2023. \\ \textbf{DOI : }\textcolor{internationalkleinblue}{\textbf{\url{10.1109/GlobConET56651.2023.10150076}}}

    \item KG Singh, \textbf{C Chodvadiya}, S Mishra, \textquotedblleft Optimizing Solar Power Forecasting Patterns using Multi-factor Cumulative Casualty matrix, \textquotedblright \,\textcolor{internationalkleinblue}{\textbf{Under Review with Professor}}

\end{enumerate}