### 2025-05-05 Title/title.tex

\begin{center}
    {{\Large \textbf{SleepGCN-Transformer: A Hybrid Graph Convolutional and
Transformer Network for Sleep Stage Classification
}}}\\
    \vspace{0.6cm}
    {\large A Master's Report}\\
    {\large Submitted in partial fulfillment of the requirement for the award of the degree of}\\
    \vspace{0.8cm}
    {\large Master of Technology in Artificial Intelligence}\\
    {\large by}\\
    \textbf{{\Large  Tanmay Rathod}}\\
    {\large 23MAI007}\\
    \vspace{0.7cm}
    {\large Under the guidance of}\\
    {\large  \textbf{Dr. Santosh Kumar Satapathy}  {\large }}\\
    {\large Department of Information and Communication Technology}\\
    \vspace{0.6cm}
        \vfill
    {\centering \includegraphics[width=0.21\textwidth]{img/pdpu_logo.png}}\\
    {\large School of Technology}\\
    {\large Pandit Deendayal Energy University}\\
    {\large Gandhinagar – 382426. Gujarat - India}\\
    {\large February, 2025}
\end{center}