### 2025-05-05 Title/abstract.tex

\begin{center}
	\section*{\textcolor{internationalkleinblue}{\textbf{{Abstract}}}}
\end{center}

Correct and effective sleep stage classification is crucial in the diagnosis of sleep disorders, conventionally done through manual interpretation of polysomnography (PSG) signals. Automated sleep stage classification is investigated in this thesis by a series of increasingly sophisticated models, from traditional machine learning to sophisticated deep learning architectures. Initial trials on the SleepEDF dataset compared models like Random Forest, XGBoost, and ensemble models, which obtained maximum accuracy of 85.35\%, and set the foundation for automated systems to perform consumer-level sleep monitoring. Future work introduced deep learning models like BiLSTM and RNNs, and BiLSTM reached 81.13\% accuracy, which proved to have better ability to model temporal sequences.

Based on these building blocks, we introduce \textit{SleepGCN-Transformer}, a new hybrid model that combines Graph Convolutional Networks (GCNs) to learn spatial relations between multi-channel EEG, EMG, and EOG signals, and Transformer encoders to learn temporal patterns. The model applies Focal Loss to handle class imbalance and a CosineAnnealingLR scheduler for dynamic learning rate adaptation. Trained with the AdamW optimizer on preprocessed 30-second epochs, SleepGCN-Transformer is 93.12\% training and 93.04\% validation accurate. Feature attribution by LIME indicates EMG and EEG Pz-Oz to be the most significant channels, demonstrating the model's interpretability.