### 2025-05-05 Title/title.tex

\begin{center}
	{{\Large \textbf{SleepGCN-Transformer: A Hybrid Graph Convolutional and\\
				Transformer Network for Sleep Stage Classification}}}\\
	\vspace{0.6cm}

	{\large A Master's Thesis}\\
	{\large Submitted in partial fulfillment of the requirement for the award of the degree of}\\
	\vspace{0.8cm}

	{\large Master of Technology in Artificial Intelligence}\\
	{\large by}\\

	\textbf{{\Large Tanmay Rathod}}\\
	{\large 23MAI007}\\
	\vspace{0.7cm}

	{\large Under the guidance of}\\
	{\large \textbf{Dr. Santosh Kumar Satapathy}}\\

	{\large Department of Information and Communication Technology}\\
	\vspace{0.6cm}

	\vfill
	{\centering \includegraphics[width=0.21\textwidth]{img/pdpu_logo.png}}\\

	{\large School of Technology}\\
	{\large Pandit Deendayal Energy University}\\
	{\large Gandhinagar – 382426. Gujarat - India}\\
	{\large May, 2025}
\end{center}