### 2025-05-08 Chapters/resources.tex

\begin{figure}[h!]
	\centering
	\begin{minipage}[b]{0.3\linewidth}
		\centering
		\includegraphics[width=\linewidth]{img/paper3}
		\caption{Automated Sleep Staging System with EEG Signal using Machine Learning Techniques}
		\label{fig:paper3}
	\end{minipage}\hfill
	\begin{minipage}[b]{0.3\linewidth}
		\centering
		\includegraphics[width=\linewidth]{img/paper1}
		\caption{Deep Neural Model for Automated Sleep Staging System using Single-Channel EEG Signal}
		\label{fig:paper1}
	\end{minipage}\hfill
	\begin{minipage}[b]{0.3\linewidth}
		\centering
		\includegraphics[width=\linewidth]{img/paper2}
		\caption{SleepGCN-Transformer: A Hybrid Graph-Convolutional and Transformer Network for Sleep Stage Classification}
		\label{fig:paper2}
	\end{minipage}
\end{figure}

\section{1. Dataset}
We conducted our experiments using the \textbf{Sleep-EDF dataset}, a publicly available dataset for sleep analysis research.\\
\texttt{https://physionet.org/content/sleep-edfx/1.0.0/}

\section{2. Development Environment}
Our implementation was carried out using the following configurations:
\begin{itemize}
	\item \textbf{Programming Language}: Python 3.13.1
	\item \textbf{Deep Learning Framework}: PyTorch
	\item \textbf{IDE}: VS Code
	\item \textbf{Hardware}: Apple M1 Chip (CPU and Metal for acceleration)
\end{itemize}

\section{3. Version Control}
To manage our codebase, we used GitHub for version control.\\
QR code available alongside the figures.

\section{4. License}
The source code of this research is registered under the \textbf{Apache License 2.0}.\\
\texttt{http://www.apache.org/licenses/}

\section{5. Key Python Libraries \& Versions}
\begin{itemize}
	\item \textbf{NumPy}: 2.2.2
	\item \textbf{Pandas}: 2.2.3
	\item \textbf{Scikit-learn}: 1.6.1
	\item \textbf{SciPy}: 1.15.1
	\item \textbf{Matplotlib}: 3.10.0
	\item \textbf{Seaborn}: 0.13.2
	\item \textbf{PyTorch}: 2.6.0
	\item \textbf{Torch Geometric}: 2.6.1
	\item \textbf{NetworkX}: 3.4.2
	\item \textbf{MNE}: 1.9.0
\end{itemize}
This setup ensures a robust and reproducible environment for the development and evaluation of our sleep stage classification model.