### 2025-05-13 Chapters/resources.tex
changes

\begin{figure}[h!]
	\centering
	\begin{minipage}[b]{0.3\linewidth}
		\centering
		\includegraphics[width=\linewidth]{img/paper3}
		\caption{Automated Sleep Staging System with EEG Signal using Machine Learning Techniques}
		\label{fig:paper3}
	\end{minipage}\hfill
	\begin{minipage}[b]{0.3\linewidth}
		\centering
		\includegraphics[width=\linewidth]{img/paper1}
		\caption{Deep Neural Model for Automated Sleep Staging System using Single-Channel EEG Signal}
		\label{fig:paper1}
	\end{minipage}\hfill
	\begin{minipage}[b]{0.3\linewidth}
		\centering
		\includegraphics[width=\linewidth]{img/paper2}
		\caption{SleepGCN-Transformer: A Hybrid Graph-Convolutional and Transformer Network for Sleep Stage Classification}
		\label{fig:paper2}
	\end{minipage}
\end{figure}

\section{Dataset}
We conducted our experiments using the \textbf{Sleep-EDF dataset}, a publicly available dataset for sleep analysis research.\\
\texttt{https://physionet.org/content/sleep-edfx/1.0.0/}

\section{Development Environment}
Our implementation was carried out using the following configurations:
\begin{itemize}
	\item \textbf{Programming Language}: Python 3.13.1
	\item \textbf{Environment Manager}: pip (with virtual environment)
	\item \textbf{Deep Learning Frameworks}: PyTorch, TensorFlow
	\item \textbf{IDE}: VS Code, Jupyter Notebook (.ipynb)
	\item \textbf{Hardware}: Apple M1 Chip (CPU and Metal for acceleration)
\end{itemize}

\section{Version Control (.GIT)}
To manage our codebase, we used GitHub for version control.\\
QR code available alongside the figures.

\section{License}
The source code of this research is registered under the \textbf{Apache License 2.0}.\\
\texttt{http://www.apache.org/licenses/}

\section{Key Python Libraries \& Versions}
\begin{itemize}
	\item \textbf{NumPy}: 2.2.2
	\item \textbf{Pandas}: 2.2.3
	\item \textbf{Scikit-learn}: 1.6.1
	\item \textbf{SciPy}: 1.15.1
	\item \textbf{Matplotlib}: 3.10.0
	\item \textbf{Seaborn}: 0.13.2
	\item \textbf{PyTorch}: 2.6.0
	\item \textbf{Torch Geometric}: 2.6.1
	\item \textbf{NetworkX}: 3.4.2
	\item \textbf{MNE}: 1.9.0
\end{itemize}
This setup ensures a robust and reproducible environment for the development and evaluation of our sleep stage classification model.

\section{published and submitted work }

\begin{itemize}
		\item \textit{Automated Sleep Staging System Using EEG Signal Feature-Based Classification by Machine Learning Techniques}, Tanmay Rathod, Santosh Kumar Satapathy, Nibedita Das, Rajesh Kumar Mohapatra, Suren Kumar Sahu, Vaishvi R Shah, 2024 IEEE 8th International Conference on Information and Communication Technology (CICT). \url{https://ieeexplore.ieee.org/document/10899578}

	\item \textit{Automated Sleep Staging System with EEG Signal using Machine Learning Techniques}, Santosh Kumar Satapathy, Tanmay Rathod, Nibedita Das, Rajesh Kumar Mohapatra, Suren Sahu, Jaynil Joshi, 2025 IEEE International Conference on Interdisciplinary Approaches in Technology and Management for Social Innovation (IATMSI). \url{https://ieeexplore.ieee.org/document/10985276}

	  \item \textit{SleepGCN-Transformer: A Hybrid Graph-Convolutional Approach for EEG-based Sleep Staging}, Tanmay Rathod et al., submitted to \textit{Biomedical Signal Processing and Control}, May 2025 (under review).

	\end{itemize}

\section{Contact Information}

For inquiries or collaboration, contact the author via the following:

\begin{itemize}
	\item \textbf{Email:} \href{mailto:tanmayrathod777@gmail.com}{tanmayrathod777@gmail.com}
	\item \textbf{GitHub:} \url{https://github.com/tanmay007thor}
	\item \textbf{IEEE Xplore Author Profile:} \url{https://ieeexplore.ieee.org/author/577823064896395}
	\item \textbf{Portfolio Website:} \url{https://tanmay-dev.netlify.app/}
	\item \textbf{LinkedIn:} \url{https://www.linkedin.com/in/tanmay-rathod-15a332230}
	\item \textbf{Phone:} +91 90165 89777
\end{itemize}