### 2025-05-19 Title/acknowledgement.tex
changes

\section*{\textcolor{internationalkleinblue}{\textbf{Acknowledgment}}}

Although this thesis bears only my name as the author, its successful completion is the result of the support, guidance, and encouragement of several individuals to whom I am deeply grateful.

First and foremost, I would like to express my sincere gratitude to my internal guide, Dr. Santosh Kumar Satapathy, from Pandit Deendayal Energy University (PDEU), for his unwavering support throughout the course of this project. His valuable insights, constructive feedback, and continuous encouragement were instrumental in shaping the direction of this research and overcoming various challenges.

I also acknowledge the faculty members and staff of the M.Tech in Artificial Intelligence program at PDEU for providing a stimulating academic environment and the foundational knowledge that enabled me to undertake this work.

Finally, I extend heartfelt thanks to my family and friends for their constant encouragement, patience, and moral support throughout this journey.

This thesis would not have been possible without the collective contributions of these individuals, and I remain truly thankful to all.

\begin{flushright}
	\textbf{Tanmay Rathod}\\
\end{flushright}