### 2025-05-19 Title/approval.tex

\begin{center}
	\textbf{\large Approval Sheet}
\end{center}

This thesis entitled \textbf{\enquote{SleepGCN-Transformer: A Hybrid Graph Convolutional and Transformer Network for Sleep Stage Classification}} by \textbf{Tanmay Rathod} is recommended for the degree of \textbf{M.Tech} in \textbf{Artificial Intelligence}.

\vspace{1cm}

\begin{table}[h!]
	\centering
	\begin{tabular}{@{}m{0.45\textwidth}<{\centering} m{0.45\textwidth}<{\centering}@{}}
		\textbf{Expert 1} & \textbf{Expert 2} \\[2.5em]
		\rule{0.35\textwidth}{0.4pt} & \rule{0.35\textwidth}{0.4pt} \\[0.5em]
		Mr. Nilesh Ranpura & Dr. Hridoy Sankar Dutta \\[0.5em]
		Director of Engineering, & Lecturer (Assistant Professor), \\[0.5em]
		Arrow Inc. (E-Infochips) & Deakin University \\[3em]
	\end{tabular}
\end{table}

\begin{table}[h!]
	\centering
	\begin{tabular}{@{}m{0.31\textwidth}<{\centering} m{0.31\textwidth}<{\centering} m{0.31\textwidth}<{\centering}@{}}
		\textbf{Guide} & \textbf{Head of Department} & \textbf{Director} \\[2.5em]
		\rule{0.25\textwidth}{0.4pt} & \rule{0.25\textwidth}{0.4pt} & \rule{0.25\textwidth}{0.4pt} \\[0.5em]
		Dr. Santosh Satapathy & Dr. Paawan Sharma & Prof.(Dr.) Dhaval Pujara \\[0.5em]
		Assistant Professor, & HoD, Dept. of ICT, & Director, \\[0.5em]
		Dept. of ICT & School of Technology & School of Technology \\
	\end{tabular}
\end{table}

\vfill
\begin{flushleft}
	Date: \makebox[1.8in]{\hrulefill}\\
	Place: \makebox[1.8in]{\hrulefill}\\    
\end{flushleft}